### 2014-11-13 tex/abstract.tex
Proofreading, fixing typos.

A fundamental challenge of using symbolic execution for software
analysis has been the treatment of dynamically allocated
data. Generalized symbolic execution (GSE) techniques have addressed
this challenge by materializing a concrete heap lazily during field
accesses.  This approach however, exacerbates the state space
explosion problem.  In this work, we present a novel symbolic
execution technique which takes inspiration from the static analysis
domain and defines a novel bipartite graph representation of the heap
and a lazy initialization algorithm that, unlike state-of-the-art
techniques, (a) supports aliasing, (b) does not require rewrite of
constraints on heap updates, (c) can initialize the heap-on-the-fly,
and (d) reduces the number of points of non-determinism during the
search. Our approach is sound and complete with respect to the
properties provable by GSE. A preliminary evaluation of our
implementation in Java Pathfinder shows the summary heap enables the
initialization of larger more complex heaps than previously possible.

>>>>

A fundamental challenge of using symbolic execution for software
analysis has been the treatment of dynamically allocated
data. Generalized symbolic execution (GSE) techniques have addressed
this challenge by materializing a concrete heap lazily during field
accesses.  This approach however, exacerbates the state space
explosion problem.  In this work, we present a novel symbolic
execution technique which takes inspiration from the static analysis
domain and defines a novel bipartite graph representation of the heap
and a lazy initialization algorithm that, unlike state-of-the-art
techniques, (a) supports aliasing, (b) does not require rewrite of
constraints on heap updates, (c) can initialize the heap-on-the-fly,
and (d) reduces the number of points of non-determinism during the
search. Our approach is sound and complete with respect to the
properties provable by GSE. A preliminary evaluation of our
implementation in Java Pathfinder shows the summary heap enables the
initialization of larger, more complex heaps than previously possible.

### 2014-11-13 tex/intro.tex
boo
Proofreading, fixing typos.

The goal of this work is to mitigate the state space explosion problem
in GSE by grouping multiple heaps together and only partitioning the
heaps at points of divergence in the control flow graph. Our
inspiration is found in the domain of static analysis that uses sets
of constraints over heap locations to encode multiple heaps in a
single representation. These sets, sometimes known as \emph{value
  sets} or \emph{value summaries}, allow multiple heaps to be
represented simultaneously with a higher degree of precision than
afforded by traditional techniques for shape analysis. Some of these
previous attempts, however, are unable to handle aliasing in heaps due
to a recursive definition of objects~\cite{..}, while others require a
set of heaps as input and are unable to initialize heaps
on-the-fly~\cite{..}.  As is typical in static analysis, over
approximation of the heap representations alleviates some of the
limitations but also leads to false positives. Also, update to heaps
require a rewrite of the constraint system making it prohitive for use
in the context of symbolic execution.

@@

In this work, to effectively represent multiple heaps simultaneously
in the context of symbolic execution we define a novel heap
representation and an algorithm to initialize and update the heap
on-the-fly. Our summary heap approach supports aliasing, it does not
require constraint rewriting on heap updates, and it reduces the
non-determinism in the search space during symbolic execution.

We represent the heap as a bipartite graph consisting of references
and locations. Each reference is able to point to multiple locations,
where each edge is predicated on constraints over aliasing between
references. Each location points to a single reference for a given
field. The use of a bipartiate graph affords two key advantages: (i)
it allows for a non-recursive definition of objects which enables us
to support aliasing, and (ii) it does not require rewriting of the
constraints during updates to the heap nor does it introduce auxillary
variables into the constraints.

This summary heap algorithm defines how the bipartite graph is updated
during lazy initialization. Unlike GSE, however, the summary heap
algorithm introduces non-determinism in the search only at points of
divergence in the control flow graph. These points of non-determinism
are at field accesses or writes that lead to null-pointer exceptions
and at comparisons of references. The former represents a divergence
due to exceptional control flow while the latter is due to program
structure.

@@

Aside from the novelty of the heap summary and its application to
symbolic execution, the new data structure appears to ameliorate state
explosion in symbolic execution as hoped. A preliminary implementation
in Java Pathfinder (JPF) shows that in general, that the heap summary
algorithm improves over other state-of-the-art techniques, and in some
instances, the improvement is remarkable: two-orders of magnitude
reduction in running time. More critically though, the heap summary
enables the initialization of larger more complex heaps than
heretofore possible as shown in our results.

@@

\item\textbf{-} A summary heap as a bipartite graph that supports
  aliasing and does not require rewriting of constraints.

@@

\item\textbf{-} A summary heap algorithm to initialize, summarize, and
  update the bipartiate graphs during symbolic execution.

@@

\item\textbf{-} A proof of concept implementation and small empirical study to demonstrate the scability of the heap summary approach
  compared to other GSE approaches.

>>>>

The goal of this work is to mitigate the state space explosion problem
in GSE by grouping multiple heaps together and only partitioning the
heaps at points of divergence in the control flow graph. Our
inspiration is found in the domain of static analysis that uses sets
of constraints over heap locations to encode multiple heaps in a
single representation. These sets, sometimes known as \emph{value
  sets} or \emph{value summaries}, allow multiple heaps to be
represented simultaneously with a higher degree of precision than
afforded by traditional techniques for shape analysis. Some of these
previous attempts, however, are unable to handle aliasing in heaps due
to a recursive definition of objects~\cite{Xie:2005}, while others require a
set of heaps as input and are unable to initialize heaps
on-the-fly~\cite{Dillig:2011,Tillmann:2008}.  As is typical in static analysis, over
approximation of the heap representations alleviates some of these
limitations but also leads to false positives. Also, updates to heaps
require a rewrite of the constraint system making it prohibtive for use
in the context of symbolic execution.

@@

In this work, to effectively represent multiple heaps simultaneously
in the context of symbolic execution we define a novel heap
representation and an algorithm to initialize and update the heap
on-the-fly. Our summary heap approach supports aliasing, does not
require constraint rewriting on heap updates, and reduces
non-determinism in the search space during symbolic execution.

We represent the heap as a bipartite graph consisting of references
and locations. Each reference is able to point to multiple locations,
where each edge is predicated on constraints over aliasing between
references. Each location points to a single reference for a given
field. The use of a bipartite graph affords two key advantages: (i)
it allows for a non-recursive definition of objects which enables us
to support aliasing, and (ii) it does not introduce auxiliary variables or 
require rewriting non-local constraints during updates to the heap.

The summary heap algorithm defines how the bipartite graph is updated
during lazy initialization. Unlike GSE, however, the summary heap
algorithm introduces non-determinism in the search only at points of
divergence in the control flow graph. These points of non-determinism
are at field accesses that lead to null-pointer exceptions
and at comparisons of references. The former represents a divergence
due to exceptional control flow while the latter is due to program
structure.

@@

%Aside from the novelty of the heap summary and its application to
%symbolic execution, the new data structure appears to ameliorate
%state explosion in symbolic execution as hoped.

A preliminary implementation in Java Pathfinder (JPF) shows that in
general, that the heap summary algorithm improves over other
state-of-the-art techniques, and in some instances, the improvement is
remarkable: two-orders of magnitude reduction in running time. More
critically though, the heap summary enables the initialization of
larger more complex heaps than heretofore possible as shown in our
results.

@@

\item\textbf{-} A symbolic summary heap representing
sets of concrete heaps without overapproximation or restricting aliasing.
% supports aliasing, and and allows strong updates without auxiliary variables 
% or constraint rewriting.

@@

\item\textbf{-} A algorithm to initialize, summarize, and
  update summary heaps during symbolic execution, without mutating 
  constraints or introducing auxiliary variables.

@@

\item\textbf{-} A proof-of-concept implementation and small empirical study 
demonstrating the scalability of the heap summary approach
  compared to other GSE approaches.
